---
title: "Challenge 1"
author: "Jeff Oliver"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

This is a data-wrangling challenge. The main goal is to produce row-wise sums
based on a suite of conditions. 

## Data: 

The data have the form:

| ID | a | b | ... | a_type      | ... |    e_type   | ... | a_1 | a_2 | ... | e_4 | e_5 |
|:---|---|---|-----|-------------|-----|-------------|-----|-----|-----|-----|-----|-----|
|  1 | 0 | 0 | ... | "in-person" | ... | "online"    | ... |  0  |  0  | ... |  1  |  0  |
|  2 | 1 | 0 | ... | "in-person" | ... | "in-person" | ... |  0  |  0  | ... |  0  |  0  |

## Goal

Create a variable which is the sum of (letter)_1 to (letter)_4, where (letter)
is a letter of the alphabet (e.g. "a", "b"). Values from each (letter) are 
_only_ included when the following conditions are true:

1. The value in the (letter) column is 1
2. The value in the (letter)_type column is "in-person"

For example, in the mock data set below, for the first row (ID = 1), values in 
`b`, `d`, and `e` columns have a value of 1 (meeting condition 1, above). Among 
those three (letters), only `e_type` has a value of "in-person" (meeting 
condition 2, above) so the sum should be computed by `e_1` + `e_2` + `e_3` + `e_4`. 

Is there a tidy way to do this?

## Data

To create the mock data set, 

```{r mock-data-creation}
mockdf <- data.frame(matrix(ncol = 36, nrow = 10))
X1 <- letters[1:5]
X2 <- paste(letters[1:5], "type", sep = "_")
X3 <- paste(rep(x = letters[1:5], each = 5), 1:5, sep = "_")
names(mockdf)<- c("ID", X1, X2, X3)
mockdf$ID <- 1:10
set.seed(1) # To ensure reproducible mock data sets
for (i in 2:6) { 
  mockdf[, i] <- sample(x = c(0,1), replace = TRUE)
}
for (i in 7:11) { 
  mockdf[, i] <- sample(x = c("in-person", "online"), replace = TRUE)
}
for (i in 12:36) { 
  mockdf[,i] = sample(x = c(0,1), replace = TRUE)
}
```

## Breadcrumbs

1. The dplyr package has several relevant data-wrangling functions. The `across` 
function may be particularly useful, in combination with `starts_with`.
2. When tackling the challenge, consider what the penultimate data might look 
like. The final data will be two-columns, with an ID and a sum. What would the 
data right _before_ this step look like? That is, what form would the data take 
to make the summation "easy."
3. While the ultimate solution may be a tidy one, getting there may require a 
trip through the valley of iteration. That is, it might take writing a `for` 
loop first before you can figure out how to do it the ~~Hadley~~ tidyverse way.
4. The first step might be to create a data set that replaces (letter)_1 through
(letter)_4 with the sum of those values; it makes for a slighly more wieldy (is 
that a word?) data set. That is, instead of a data set with columns `a_1`, 
`a_2`, ..., `a_5`, it only has `a_sum = sum(a_1, a_2, a_3, a_4)`. Whether or 
not the sum is used can be a decision further downstream. For a single (letter), 
this can be done via
`mockdf %>% rowwise() %>% mutate(a_sum = sum(across(num_range(prefix = "a_", range = 1:4))))`
Employing regular expressions might make it possible to do this for all 
(letter)s...
5. A word of caution: be sure to build-in cases where _none_ of the (letter)s 
meet both criteria (coder's choice as to what to do in those cases (`0` or 
`NA`))