---
title: "Challenge 1"
author: "Kelsey Gonzalez"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
```

# Description

This is a data-wrangling challenge. The main goal is to produce row-wise sums
based on a suite of conditions. 

## Data: 

The data have the form:

| ID | a | b | ... | a_type      | ... |    e_type   | ... | a_1 | a_2 | ... | e_4 | e_5 |
|:---|---|---|-----|-------------|-----|-------------|-----|-----|-----|-----|-----|-----|
|  1 | 0 | 0 | ... | "in-person" | ... | "online"    | ... |  0  |  0  | ... |  1  |  0  |
|  2 | 1 | 0 | ... | "in-person" | ... | "in-person" | ... |  0  |  0  | ... |  0  |  0  |

## Goal

Create a variable which is the sum of (letter)_1 to (letter)_4, where (letter)
is a letter of the alphabet (e.g. "a", "b"). Values from each (letter) are 
_only_ included when the following conditions are true:

1. The value in the (letter) column is 1
2. The value in the (letter)_type column is "in-person"

For example, in the mock data set below, for the first row (ID = 1), values in 
`c` and `e` columns have a value of 1 (meeting condition 1, above). Between 
those two (letters), only `c_type` has a value of "in-person" (meeting 
condition 2, above) so the sum should be computed by `c_1` + `c_2` + `c_3` + `c_4`. 

Is there a tidy way to do this?

## Data

To create the mock data set, 

```{r mock-data-creation}
mockdf <- data.frame(matrix(ncol = 36, nrow = 10))
X1 <- letters[1:5]
X2 <- paste(letters[1:5], "type", sep = "_")
X3 <- paste(rep(x = letters[1:5], each = 5), 1:5, sep = "_")
names(mockdf)<- c("ID", X1, X2, X3)
mockdf$ID <- 1:10
set.seed(1) # To ensure reproducible mock data sets
for (i in 2:6) { 
  mockdf[, i] <- sample(x = c(0,1), replace = TRUE)
}
for (i in 7:11) { 
  mockdf[, i] <- sample(x = c("in-person", "online"), replace = TRUE)
}
for (i in 12:36) { 
  mockdf[,i] = sample(x = c(0,1), replace = TRUE)
}
```

# Solution

Provided by [Kelsey Gonzalez](https://github.com/kelseygonzalez)

Start: 

| ID | a | b | ... | a_type      | ... |    e_type   | ... | a_1 | a_2 | ... | e_4 | e_5 |
|:---|---|---|-----|-------------|-----|-------------|-----|-----|-----|-----|-----|-----|
|  1 | 0 | 0 | ... | "in-person" | ... | "online"    | ... |  0  |  0  | ... |  1  |  1  |
|  2 | 1 | 0 | ... | "in-person" | ... | "in-person" | ... |  0  |  0  | ... |  0  |  1  |

_Note_: The table above reflects the _shape_ of the data at this point; actual 
data values may differ.

## Step 1

Pivot data so a row includes data from a single (letter); assuming value in 
each (letter) column refers to the tie strength (and names column accordingly).

```{r pivot-alters}
midpoint <- mockdf %>% 
  pivot_longer(cols = c(a_type:e_type), names_to = "alter_type", values_to = "type") %>% 
  pivot_longer(cols = c(a:e), names_to = "alter", values_to = "tie_strength") %>% 
  pivot_longer(cols = c(a_1:e_5), names_pattern = "(.)_(.)", names_to = c("alter_q", "question")) %>%  
  pivot_wider(id_cols = c(ID, alter_type, type, alter,tie_strength, alter_q), 
              names_from = "question",  names_glue = "q{question}",
              values_from = "value") %>% 

  mutate(alter_type = str_replace(alter_type, "_type", "")) %>% 
    filter(alter == alter_q ,
           alter == alter_type)  %>% 
  select(ID, alter, tie_strength, type, q1:q5)

midpoint
```


Mid1:

| ID | Alter | type        | ... |    type     | ... | q1  | q2  | q3  | q4  | q5  |
|:---|-------|-------------|-----|-------------|-----|-----|-----|-----|-----|-----|
|  1 | a     | "in-person" | ... | "online"    | ... |  0  |  0  |  0  |  1  |  0  |
|  2 | b     | "in-person" | ... | "in-person" | ... |  0  |  0  |  0  |  1  |  1  |

_Note_: The table above reflects the _shape_ of the data at this point; actual 
data values may differ.

## Step 2

Perform sums based on two conditions described in challenge (tie strength 
equals 1 and type is "in-person")

```{r sum-alters}
alter_sums <- midpoint %>%  
  filter(type == "in-person",
         tie_strength == 1) %>% 
  select(-type) %>% 
  rowwise(ID) %>% 
  mutate(sum = sum(c_across(num_range(prefix = "q", range = 1:4)))) %>%
  select(-starts_with("q"))

alter_sums
```

Per-alter sums:

| ID | alter | tie_strength | sum | 
|:---|-------|--------------|-----|
|  1 | c     | 1            | 3   | 
|  2 | a     | 1            | 0   | 
|  2 | e     | 1            | 3   | 

_Note_: The table above reflects the _shape_ of the data at this point; actual 
data values may differ.

## Step 3

Sum per-alter values for each ID.

```{r sum-ids}
final_sums <- alter_sums %>%
  group_by(ID) %>%
  summarize(sum = sum(sum)) # My apologies. jcoliver is to blame

final_sums
```

Final sums for each ID:

| ID | sum | 
|:---|-----|
|  1 | 3   | 
|  2 | 3   | 
|  3 | 3   | 
